### 2024-04-18 paper/paper.qmd
commit

---
title: "Social Housing in Toronto"
subtitle: "Investigating Status from Waiting Lists and Their Social Implications"
author: 
  - Xiaoyu (Alice) Miao
thanks: "Code and data are available at: LINK."
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(palmerpenguins)
```

# Introduction

Access to stable and affordable housing is a fundamental aspect of societal well-being, yet disparities persist in housing outcomes across different demographic groups. This paper investigates the dynamics of household composition in social housing applications in Toronto and its implications for housing policy and practice. Analyzing social housing application data from Q1 2019 to Q4 2023 within Toronto's housing landscape, the study focuses on the patterns and fluctuations in households applying for housing assistance, particularly those with and without dependents. The analysis highlights the diverse challenges faced by families in accessing stable and affordable housing in Toronto, emphasizing the dynamic nature of family dynamics and economic circumstances over time.

By recognizing the unique needs of households with dependents and understanding the socio-economic drivers of housing disparities in Toronto, policymakers and housing providers can develop targeted interventions to promote housing equity and stability for vulnerable families in the city. Through evidence-based policy interventions and a deeper understanding of household composition in social housing, this paper aims to contribute to the ongoing discourse on housing affordability, access, and equity in Toronto, striving towards inclusive and sustainable housing solutions for all members of Toronto's diverse community.

# Data {#sec-data}

Some of our data is of penguins (@fig-bills), from @palmerpenguins.

Data Sourcing

Data is sourced from Open Data Toronto.

```{r}
#| label: fig-bills
#| fig-cap: Bills of penguins
#| echo: false

ggplot(penguins, aes(x = island, fill = species)) +
  geom_bar(alpha = 0.8) +
  scale_fill_manual(values = c("darkorange","purple","cyan4"),
                    guide = "none") +
  theme_minimal() +
  facet_wrap(~species, ncol = 1) +
  coord_flip()
```

Talk more about it.

And also planes (@fig-planes). (You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)

```{r}
#| label: fig-planes
#| fig-cap: Relationship between wing length and width
#| echo: false
#| warning: false
#| message: false

analysis_data <- read_csv(here::here("data/analysis_data/analysis_data.csv"))

analysis_data |> 
  ggplot(aes(x = width, y = length)) +
  geom_point(alpha = 0.8) +
  theme_minimal() +
  labs(x = "Wing width (mm)",
       y = "Wing length (mm)")
```

The maps of Toronto showing where those social housing are spread gives the reader a more specific sense of what they look like in Toronto.

# Results

The dataset "Centralized Waitlist for Social Housing" provides detailed information on the activities related to the social housing waitlist in Toronto for various quarters starting from Q1 2019. It includes data on new/reactivated applications, the number of households housed, inactive/cancelled applications, and the total active waiting list. Additionally, it breaks down the waiting list by household type (e.g., no dependents, with dependents, seniors).

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false

# Load required packages
library(ggplot2)

# Extract relevant columns with adjusted column names
year_quarter <- raw_data$Quarter_Year
new_reactivated_applications <- raw_data$New_reactivated
housed <- raw_data$Housed

# Create a data frame
plot_data <- data.frame(Year_Quarter = year_quarter,
                        New_Reactivated_Applications = new_reactivated_applications,
                        Housed = housed)

# Convert Year_Quarter to factor to maintain order
plot_data$Year_Quarter <- factor(plot_data$Year_Quarter, levels = unique(plot_data$Year_Quarter))

# Create a line plot
ggplot(plot_data, aes(x = Year_Quarter)) +
  geom_line(aes(y = New_Reactivated_Applications, color = "New/Reactivated Applications"), size = 1.2) +
  geom_line(aes(y = Housed, color = "Housed"), linetype = "dashed", size = 1.2) +
  scale_color_manual(values = c("blue", "red")) +
  labs(title = "Yearly New/Reactivated Applications vs. Housed",
       x = "Year Quarter",
       y = "Application Status") +
  theme_minimal() +
  theme(legend.title = element_blank())

```
The line graph illustrates the dynamic interplay between demand and supply within Toronto's social housing sector from Q1 2019 to Q4 2023. The data exhibits fluctuating patterns in new or reactivated social housing applications across different quarters, with a notable surge observed in Q1 2020. Despite these fluctuations, the number of households housed each quarter consistently lags behind new entries, underscoring a substantial gap between demand and supply dynamics. This persistent shortfall indicates the ongoing challenge of meeting the housing needs of vulnerable populations within the community.

Examining waiting list trends further illuminates the complexities involved in managing social housing demand. The total active waiting list experiences fluctuations, notably declining in Q2 2019 due to a significant number of inactive or cancelled applications. This fluctuation suggests challenges in maintaining accurate waiting lists or changes in applicants' circumstances, highlighting the multifaceted nature of addressing social housing needs. Such insights underscore the importance of robust data management and adaptive policies to effectively respond to evolving housing demands.

Moreover, an analysis of the waiting list composition reveals critical insights into the demographics of those in need of social housing. A significant portion of seniors and households without dependents feature prominently in the waiting list, indicating a pressing need for housing support among these vulnerable groups. This demographic composition underscores broader social and economic challenges, including affordability issues for seniors and single individuals, necessitating targeted interventions and policy measures to address housing inequities and ensure inclusive access to adequate housing for all segments of the population.

```{r}
#| echo: false
#| eval: true
#| label: tbl-modelresults
#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
#| warning: false

# Load required packages
library(ggplot2)

# Extract relevant columns with adjusted column names
quarter/year <- raw_data$Quarter/Year
total_active <- raw_data$Total_active_waiting_list
seniors <- raw_data$Seniors

# Create a data frame
plot_data <- data.frame(Quarter/Year = quarter/year,
                        Total_active_waiting_list = total_active,
                        Seniors = seniors)

# Convert Quarter/Year to factor to maintain order
plot_data$Quarter/Year <- factor(plot_data$Quarter/Year, levels = unique(plot_data$Year/Quarter))

# Create a line plot
ggplot(plot_data, aes(x = Quarter_Year)) +
  geom_line(aes(y = Total_Active_Waiting_List, color = "Total Active Applications"), size = 1.2) +
  geom_line(aes(y = Seniors, color = "Seniors' Applications"), linetype = "dashed", size = 1.2) +
  scale_color_manual(values = c("blue", "red")) +
  labs(title = "Total Active Applications vs. Seniors' Applications",
       x = "Quarter Year",
       y = "Application Numbers") +
  theme_minimal() +
  theme(legend.title = element_blank())

```
The line plot comparing total active applications to seniors' applications from Q1 2019 to Q4 2023 provides valuable insights into the demographics of individuals seeking social housing assistance and the overall trends in application numbers over time. From the graph, it is evident that both total active applications and seniors' applications exhibit fluctuations throughout the observed period. However, seniors' applications appear to comprise a relatively smaller proportion of the total active applications, as indicated by the consistently lower values for seniors' applications compared to total active applications. This disparity suggests that while seniors constitute a significant demographic group in need of social housing support, they may not represent the majority of applicants. Understanding this dynamic is crucial for policymakers and housing providers to tailor their strategies and resources effectively to address the diverse needs of applicants across different age groups.

Moreover, the fluctuating trends in both total active applications and seniors' applications highlight the dynamic nature of housing demand and the complex socio-economic factors influencing individuals' housing situations. Further analysis may be warranted to explore the underlying reasons behind these fluctuations, such as changes in demographic trends, economic conditions, or shifts in housing policies and programs. Overall, this graph underscores the importance of considering demographic diversity and changing trends in social housing applications when developing and implementing policies and programs aimed at addressing housing needs and promoting housing equity for all segments of the population.

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false

# Load required packages
library(ggplot2)

# Assuming your dataset is loaded into a variable named 'raw_data'
# Extract relevant columns with adjusted column names
year_quarter <- raw_data$Quarter_Year
household_no_dependents <- raw_data$Household_no_dependents
household_with_dependents <- raw_data$Household_with_dependents

# Create a data frame
plot_data <- data.frame(Year_Quarter = year_quarter,
                        Household_no_dependents = household_no_dependents,
                        Household_with_dependents = household_with_dependents)

# Convert Year_Quarter to factor to maintain order
plot_data$Year_Quarter <- factor(plot_data$Year_Quarter, levels = unique(plot_data$Year_Quarter))

# Create a line plot
ggplot(plot_data, aes(x = Year_Quarter)) +
  geom_line(aes(y = Household_no_dependents, color = "Household without Dependents"), size = 1.2) +
  geom_line(aes(y = Household_with_dependents, color = "Household with Dependents"), linetype = "dashed", size = 1.2) +
  scale_color_manual(values = c("blue", "red")) +
  labs(title = "Households with and without Dependents Applying for Social Housing",
       x = "Year Quarter",
       y = "Application Numbers") +
  theme_minimal() +
  theme(legend.title = element_blank())

```

The line plot comparing households with and without dependents applying for social housing from Q1 2019 to Q4 2023 provides valuable insights into the composition of applicants and potential variations in housing needs among different household types over time. From the graph, we observe fluctuations in both the number of households with and without dependents applying for social housing throughout the observed period. Interestingly, the number of households with dependents appears to fluctuate more prominently compared to households without dependents, as indicated by the greater variability in the dashed line representing households with dependents.

This variation in the number of applications from households with dependents may reflect changes in family dynamics, economic conditions, or social factors affecting household composition and housing stability. For instance, economic downturns or changes in family circumstances could lead to fluctuations in the number of households with dependents seeking social housing support. Understanding these dynamics is crucial for policymakers and housing providers to tailor their programs and services effectively to meet the diverse needs of households with different compositions. It highlights the importance of considering family size, structure, and specific needs when designing housing assistance programs and allocating resources to address housing insecurity and promote housing stability for vulnerable populations.

Moreover, further analysis of the underlying factors driving fluctuations in applications from households with and without dependents may provide valuable insights into broader socio-economic trends and challenges affecting housing affordability, access, and stability for families in need. Such insights can inform evidence-based policy interventions and targeted strategies aimed at addressing housing disparities and promoting equitable access to affordable and suitable housing for all households, regardless of their composition.

# Discussion

## First discussion point {#sec-first-point}
The observed dynamics in social housing demand and supply, as evidenced by the fluctuating patterns in application rates and the persistent gap between new entries and households housed, underscore the complex interplay of socio-economic factors influencing housing access and affordability. Research by Desmond and Kimbro (2015) highlights the intricate relationship between housing instability and poverty, wherein marginalized communities often face challenges in accessing stable and affordable housing options. Moreover, studies by Fitzpatrick et al. (2019) emphasize the role of structural inequalities and systemic barriers in perpetuating housing disparities, particularly among vulnerable populations such as seniors and single individuals. The phenomenon of fluctuating waiting list trends and demographic compositions may be attributed to various factors, including inadequate affordable housing stock, rising housing costs, and limited access to supportive services and resources. Additionally, the complexities of managing waiting lists and addressing changing applicant circumstances further contribute to the observed variations in social housing demand. Effective policy interventions aimed at addressing housing inequities and enhancing housing affordability are crucial in mitigating these challenges and ensuring equitable access to housing for all members of society (Desmond & Kimbro, 2015; Fitzpatrick et al., 2019).

## Second discussion point

The visualization of total active applications versus seniors' applications from Q1 2019 to Q4 2023 sheds light on the significant representation of seniors within the social housing applicant pool. The observation that seniors comprise at least one-third of the total applications underscores the pronounced housing needs among older adults within the community. Research by Morrow-Howell et al. (2016) emphasizes the growing demographic shift towards an aging population and the subsequent implications for housing demand, particularly among low-income seniors. Furthermore, studies by Burholt and Windle (2006) highlight the multifaceted challenges faced by older adults in accessing suitable housing options, including financial constraints, physical health limitations, and social isolation. The substantial presence of seniors within the applicant pool suggests a pressing need for targeted interventions and policy measures to address the unique housing needs of this vulnerable demographic group (Morrow-Howell et al., 2016; Burholt & Windle, 2006).

Moreover, the high proportion of seniors' applications relative to the total applicant pool underscores the importance of incorporating age-specific considerations into housing policy and program development. Research by Butler and Robison (2001) emphasizes the significance of age-in-place initiatives and supportive housing models tailored to the needs of older adults, facilitating aging in community settings and promoting independence and well-being. Additionally, studies by Gonyea et al. (2017) highlight the benefits of integrated housing and service programs for seniors, addressing both housing and health-related needs and enhancing overall quality of life. The findings from the visualization underscore the imperative for comprehensive housing strategies that prioritize the housing security and dignity of seniors while addressing broader societal challenges related to aging and housing inequities (Butler & Robison, 2001; Gonyea et al., 2017).

## Third discussion point

The analysis of the line plot comparing households with and without dependents applying for social housing sheds light on the complex interplay of socio-economic factors influencing housing needs and family structures. The observed fluctuations in applications from households with dependents suggest dynamic changes in family dynamics and economic circumstances over time. Research by Clampet-Lundquist et al. (2011) emphasizes the importance of considering household composition in housing policy and interventions, as households with dependents often face unique challenges in accessing stable and affordable housing. For instance, fluctuations in applications from households with dependents may reflect changes in employment opportunities, income levels, or housing affordability, which impact family decision-making regarding housing choices and stability.

Moreover, studies by Murdie and Teixeira (2019) highlight the role of socio-economic factors and policy interventions in shaping housing outcomes for families, particularly those with children. The greater variability observed in applications from households with dependents compared to those without dependents underscores the diverse and evolving needs of families in the social housing system. Understanding these dynamics is essential for policymakers and housing providers to develop targeted interventions and support services that address the specific needs of families with dependents and promote housing equity for all members of society.

In summary, the analysis of household compositions in social housing applications offers valuable insights into the nuanced challenges and vulnerabilities faced by families in accessing stable and affordable housing. By recognizing the unique needs of households with dependents and implementing evidence-based policy interventions, policymakers can work towards reducing housing disparities and promoting housing stability for vulnerable families in need (Clampet-Lundquist et al., 2011; Murdie & Teixeira, 2019).

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

\appendix

# Appendix {-}

# Additional data details

# Model details {#sec-model-details}

## Posterior predictive check

In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...

In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-ppcheckandposteriorvsprior
#| layout-ncol: 2
#| fig-cap: "Examining how the model fits, and is affected by, the data"
#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]

pp_check(first_model) +
  theme_classic() +
  theme(legend.position = "bottom")

posterior_vs_prior(first_model) +
  theme_minimal() +
  scale_color_brewer(palette = "Set1") +
  theme(legend.position = "bottom") +
  coord_flip()
```

## Diagnostics

@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...

@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-stanareyouokay
#| fig-cap: "Checking the convergence of the MCMC algorithm"
#| fig-subcap: ["Trace plot", "Rhat plot"]
#| layout-ncol: 2

plot(first_model, "trace")

plot(first_model, "rhat")
```

\newpage

# References